# Results

### Continent-Wise CO<sub>2</sub> Emissions Over 3 Decades
```{r}
#| warning: false

library(countrycode)
library(dplyr)
library(ggplot2)
library(forcats)

emissions_df = read.csv("emission_analysis.csv")

emissions_df_g1 <- emissions_df

# For this purpose, we are filling all missing CO2 emission values with 0
emissions_df_g1$CO2_Emissions[is.na(emissions_df_g1$CO2_Emissions)] <- 0

# Obtaining Continents
emissions_df_g1$Continent <- countrycode(sourcevar = emissions_df_g1[, "Country.Code"],
                           origin = "iso3c",
                           destination = "continent")

# Adding decade ranges and aggregating CO2 emissions
emissions_df_g1 <- emissions_df_g1 |>
  mutate(
    Year_Range = case_when(
      Year >= 1990 & Year <= 2000 ~ "1990-2000",
      Year >= 2000 & Year <= 2010 ~ "2000-2010",
      Year >= 2010 & Year <= 2020 ~ "2010-2020")) |>
  group_by(Continent, Year_Range) |>                
  summarise(Aggregate_CO2 = sum(CO2_Emissions)) |>
  ungroup()

ggplot(emissions_df_g1, aes(x = Aggregate_CO2, y = fct_reorder(Continent, Aggregate_CO2))) +
  geom_col(fill = "cornflowerblue") +
  facet_wrap(~Year_Range) +
  labs(title = expression("Continent-Wise CO"[2] ~ "Emissions Over 3 Decades"), x = "Aggregate CO2 Emissions (megatonnes)", y = "Continents")

```
From the graph above we can study the amount and distribution of CO<sub>2</sub> emissions over 3 decades across different continents.
In all facets, we observe the same order of CO<sub>2</sub> emission amounts with Asia having the highest amount due to the continent have the most number of countries, and Oceania (Australia and New Zealand) having the least amount.
The amount of CO<sub>2</sub> emissions in Asia steadily increased each decade and reached an alarming 200k megatonnes, which indicates that Asia plays a huge role in pollution.
However, the amount of CO<sub>2</sub> emissions in Europe decreased over each decade, having its highest emissions in 1990-2000. This shows a slow down in CO<sub>2</sub> emission activity in this continent.
The amount of CO<sub>2</sub> emissions in Africa slightly increased each decade, while they remained almost the same in the Americas and Oceania from 1990-2020.

### Biplot to understand relationships between contributing factors
```{r}
#| warning: false
#| fig.height: 8
library(redav)

emissions_df_g2 <- emissions_df
emissions_df_g2[is.na(emissions_df)] <- 0
draw_biplot(emissions_df_g2, points = FALSE) +
  labs(title = "Biplot of CO"[2] ~ "Emissions Data")
```
Using the biplot above we can analyze the relationship of different features that contribute to CO<sub>2</sub> emissions. We can observe that Imports, Exports and GDP are highly correlated to the amount of CO<sub>2</sub> emissions, while Fossil Fuel Consumption shows a positive correlation. Livestock Production is almost at a 90 degrees angle to the CO<sub>2</sub> emissions vector and therefore we can say that there is no correlation between the two. Features such as Air Pollution and Renewable Energy have a negative correlation. Air pollution having a negative correlation could be due to the fact that air pollutants are made up of more than just CO<sub>2</sub>, air pollutants also include sulphur dioxide, nitrous oxide, carbon monoxide, etc. which have significantly reduced over the years.

### Analyzing the Trend of CO<sub>2</sub> Emissions and Forest Area over 30 Years
```{r}

# Countries in the Amazon Rainforest: Brazil, Peru, Bolivia, Ecuador, Colombia, Venezuela, French Guiana, Guyana, Suriname
# Source: https://www.nature.org/en-us/get-involved/how-to-help/places-we-protect/amazon-rainforest/

amazon_countries <- c("BRA", "PER", "BOL", "ECU", "COL", "VEN", "GUF", "GUY", "SUR")

emissions_df_amazon <- emissions_df|> 
  filter(Country.Code %in% amazon_countries)

# We do not have data for French Guiana so we will go ahead and account for the rest of the 8 countries in the Amazon Rainforest
# The filtered dataset does not have missing values for Forest Area and CO2 Emissions

# aggregating CO2 emissions and Forest Area by Year
aggregated_data <- emissions_df_amazon |>
  group_by(Year) |>
  summarize(
    Total_CO2 = sum(CO2_Emissions),
    Total_Forest_Area = sum(Forest_Area)
  )

# Scaling the data
aggregated_data <- aggregated_data |>
  mutate(
    Scaled_CO2 = (Total_CO2 - min(Total_CO2)) / (max(Total_CO2) - min(Total_CO2)),
    Scaled_Forest_Area = (Total_Forest_Area - min(Total_Forest_Area)) / (max(Total_Forest_Area) - min(Total_Forest_Area))
  )

ggplot(aggregated_data, aes(x = Year)) +
  geom_line(aes(y = Scaled_CO2, color = "CO2 Emissions (Mt)")) +
  geom_line(aes(y = Scaled_Forest_Area, color = "Forest Area (%)")) +
  scale_color_manual(values = c("CO2 Emissions (Mt)" = "red", "Forest Area (%)" = "#228B22")) +
  labs(
    title = "CO"[2] ~ "Emissions and Forest Area Trend in the Amazon Rainforest (1990-2020)",
    x = "Year",
    y = "Scaled Values",
    color = "Legend"
  ) +
  theme(legend.position = "right")

```
Through this graph we wanted to investigate how the Amazon Rainforest has been impacted by deforestation and other human activities, and if this has impacted the CO<sub>2</sub> emissions in the rainforest. Using the visualization above, we notice a clear trend of forest area dramatically reducing and CO<sub>2</sub> emissions steadily increasing (with the exception of years 2015-2020). This confirms the fact that the depletion of the Amazon rainforest, one of the largest carbon sinks in the world has caused an increase in CO<sub>2</sub> emissions, with the highest emissions being released in 2014.

### Determining if Livestock Production Contributes to CO<sub>2</sub> emissions
The effect of livestock production on CO<sub>2</sub> emissions is currently being studied and debated. We have investigated this topic using the graph below
Source: https://sentientmedia.org/how-does-livestock-affect-climate-change/
```{r}

# Aggregating CO2 Emissions and Livestock Production by Country
emissions_df_livestock <- emissions_df |>
  group_by(Country.Name) |>
  summarize(
    Total_CO2 = sum(CO2_Emissions),
    Total_Livestock = sum(Livestock_Production)
  )

emissions_df_livestock <- na.omit(emissions_df_livestock)

# Obtaining the top 3 and bottom 3 countries in accordance with livestock production
sorted_emissions <- emissions_df_livestock |>
  arrange(desc(Total_Livestock))

top_3_countries <- sorted_emissions |>
  slice_head(n = 3)

bottom_3_countries <- sorted_emissions |>
  slice_tail(n = 3)

combined_df <- bind_rows(top_3_countries |> mutate(Group = "Countries with Highest Production"),
                         bottom_3_countries |> mutate(Group = "Countries with Lowest Production"))

ggplot(combined_df, aes(x = Total_Livestock, y = Total_CO2, color = Group)) +
  geom_point(size = 3) +
  geom_text(aes(label = Country.Name), vjust = -1, size = 3, show.legend = FALSE) +
  scale_color_manual(values = c("Countries with Highest Production" = "darkgreen", "Countries with Lowest Production" = "darkorange")) +
  labs(
    title = "CO"[2] ~ "Emissions over Countries with Lowest and Highest Livestock Production",
    x = "Total Livestock Production",
    y = "Total CO2 Emissions (Mt)",
    color = "Group"
  ) +
  theme(legend.position = "bottom")

```
We have visualized the emissions of the countries with lowest and highest livestock production and observe that there is no clear relationship between CO<sub>2</sub> emissions and livestock production. Countries with the highest livestock production (St. Kitts and Nevis & Antigua and Barbuda) have the lowest emissions. Countries with the lowest livestock production have higher emissions. Kuwait which has livestock production on the lower end is observed to have the highest emissions. The absence of a clear correlation could be because livestock production activities mainly release a substantial amount of methane as a pollutant as compared to CO<sub>2</sub>.

### Analyzing the Trend of CO<sub>2</sub> Emissions and Fossil Fuel Consumption over 30 Years
```{r}
emissions_df_fuel <- emissions_df

emissions_df_fuel <- emissions_df_fuel[complete.cases(emissions_df_fuel[c("CO2_Emissions", "Fossil_Fuel_Energy")]), ]

aggregated_data <- emissions_df_fuel |>
  group_by(Year) |>
  summarize(
    Total_CO2 = sum(CO2_Emissions),
    Total_Fuel = sum(Fossil_Fuel_Energy)
  )

aggregated_data <- aggregated_data |>
  mutate(
    Scaled_CO2 = (Total_CO2 - min(Total_CO2)) / (max(Total_CO2) - min(Total_CO2)),
    Scaled_Fuel = (Total_Fuel - min(Total_Fuel)) / (max(Total_Fuel) - min(Total_Fuel))
  )

ggplot(aggregated_data, aes(x = Year)) +
  geom_line(aes(y = Scaled_CO2, color = "CO2 Emissions (Mt)")) +
  geom_line(aes(y = Scaled_Fuel, color = "Fossil Fuel Energy Consumption (%)")) +
  scale_color_manual(values = c("CO2 Emissions (Mt)" = "red", "Fossil Fuel Energy Consumption (%)" = "black")) +
  labs(
    title = "Worldwide CO"[2] ~ "Emissions and Fossil Fuel Consumption Trend (1990-2015)",
    x = "Year",
    y = "Scaled Values",
    color = "Legend"
  ) +
  theme(legend.position = "right")

```
The values for 2016-2020 are missing for fossil fuel consumption, hence this analysis is done for the years 1990-2015. From the graph above, we can see the fossil fuel consumption remains almost constant throughout the years (on the higher end) and the CO<sub>2</sub> emissions steadily increase reflecting the rising global energy demand. Fossil fuels remain as the primary source of energy from 1990 to present and have contributed to increase in global CO<sub>2</sub> emissions. The drop in values in 2015 is due to scaling and represents that both CO<sub>2</sub> emissions and fossil fuel consumption were the lowest in this year.
