# Results

Continent-Wise CO<sub>2</sub> Emissions Over 3 Decades
```{r}
#| warning: false

library(countrycode)
library(dplyr)
library(ggplot2)
library(forcats)

emissions_df = read.csv("emission_analysis.csv")

emissions_df_g1 <- emissions_df

# For this purpose, we are filling all missing CO2 emission values with 0
emissions_df_g1$CO2_Emissions[is.na(emissions_df_g1$CO2_Emissions)] <- 0

# Obtaining Continents
emissions_df_g1$Continent <- countrycode(sourcevar = emissions_df_g1[, "Country.Code"],
                           origin = "iso3c",
                           destination = "continent")

# Adding decade ranges and aggregating CO2 emissions
emissions_df_g1 <- emissions_df_g1 |>
  mutate(
    Year_Range = case_when(
      Year >= 1990 & Year <= 2000 ~ "1990-2000",
      Year >= 2000 & Year <= 2010 ~ "2000-2010",
      Year >= 2010 & Year <= 2020 ~ "2010-2020")) |>
  group_by(Continent, Year_Range) |>                
  summarise(Aggregate_CO2 = sum(CO2_Emissions)) |>
  ungroup()

ggplot(emissions_df_g1, aes(x = Aggregate_CO2, y = fct_reorder(Continent, Aggregate_CO2))) +
  geom_col(fill = "cornflowerblue") +
  facet_wrap(~Year_Range) +
  labs(title = expression("Continent-Wise CO"[2] ~ "Emissions Over 3 Decades"), x = "Aggregate CO2 Emissions (megatonnes)", y = "Continents")

```
From the graph above we can study the amount and distribution of CO<sub>2</sub> emissions over 3 decades across different continents.
In all facets, we observe the same order of CO<sub>2</sub> emission amounts with Asia having the highest amount due to the continent have the most number of countries, and Oceania (Australia and New Zealand) having the least amount.
The amount of CO<sub>2</sub> emissions in Asia steadily increased each decade and reached an alarming 200k megatonnes, which indicates that Asia plays a huge role in pollution.
However, the amount of CO<sub>2</sub> emissions in Europe decreased over each decade, having its highest emissions in 1990-2000. This shows a slow down in CO<sub>2</sub> emission activity in this continent.
The amount of CO<sub>2</sub> emissions in Africa slightly increased each decade, while they remained almost the same in the Americas and Oceania from 1990-2020.

Creating a Biplot to understand relationships between contributing factors
```{r}
#| warning: false
#| fig.height: 8
library(redav)

emissions_df_g2 <- emissions_df
emissions_df_g2[is.na(emissions_df)] <- 0
draw_biplot(emissions_df_g2, points = FALSE) +
  labs(title = "Biplot of CO"[2] ~ "Emissions Data")
```

Analyzing the Trend of CO<sub>2</sub> Emissions and Forest Area over 30 Years
```{r}

# Countries in the Amazon Rainforest: Brazil, Peru, Bolivia, Ecuador, Colombia, Venezuela, French Guiana, Guyana, Suriname
# Source: https://www.nature.org/en-us/get-involved/how-to-help/places-we-protect/amazon-rainforest/

amazon_countries <- c("BRA", "PER", "BOL", "ECU", "COL", "VEN", "GUF", "GUY", "SUR")

emissions_df_amazon <- emissions_df|> 
  filter(Country.Code %in% amazon_countries)

# We do not have data for French Guiana so we will go ahead and account for the rest of the 8 countries in the Amazon Rainforest
# The filtered dataset does not have missing values for Forest Area and CO2 Emissions

# aggregating CO2 emissions and Forest Area by Year
aggregated_data <- emissions_df_amazon |>
  group_by(Year) |>
  summarize(
    Total_CO2 = sum(CO2_Emissions),
    Total_Forest_Area = sum(Forest_Area)
  )

# Scaling the data
aggregated_data <- aggregated_data |>
  mutate(
    Scaled_CO2 = (Total_CO2 - min(Total_CO2)) / (max(Total_CO2) - min(Total_CO2)),
    Scaled_Forest_Area = (Total_Forest_Area - min(Total_Forest_Area)) / (max(Total_Forest_Area) - min(Total_Forest_Area))
  )

ggplot(aggregated_data, aes(x = Year)) +
  geom_line(aes(y = Scaled_CO2, color = "CO2 Emissions (Mt)")) +
  geom_line(aes(y = Scaled_Forest_Area, color = "Forest Area (%)")) +
  scale_color_manual(values = c("CO2 Emissions (Mt)" = "red", "Forest Area (%)" = "#228B22")) +
  labs(
    title = "CO"[2] ~ "Emissions and Forest Area Trend in the Amazon Rainforest (1990-2020)",
    x = "Year",
    y = "Scaled Values",
    color = "Legend"
  ) +
  theme(legend.position = "right")

```

Determining if Livestock Production Contributes to CO<sub>2</sub> emissions
```{r}
# It is being studied that livestock production increases greenhouse gas emissions
# Source: https://sentientmedia.org/how-does-livestock-affect-climate-change/

# Aggregating CO2 Emissions and Livestock Production by Country
emissions_df_livestock <- emissions_df |>
  group_by(Country.Name) |>
  summarize(
    Total_CO2 = sum(CO2_Emissions),
    Total_Livestock = sum(Livestock_Production)
  )

emissions_df_livestock <- na.omit(emissions_df_livestock)

# Obtaining the top 3 and bottom 3 countries in accordance with livestock production
sorted_emissions <- emissions_df_livestock |>
  arrange(desc(Total_Livestock))

top_3_countries <- sorted_emissions |>
  slice_head(n = 3)

bottom_3_countries <- sorted_emissions |>
  slice_tail(n = 3)

combined_df <- bind_rows(top_3_countries |> mutate(Group = "Countries with Highest Production"),
                         bottom_3_countries |> mutate(Group = "Countries with Lowest Production"))

ggplot(combined_df, aes(x = Total_Livestock, y = Total_CO2, color = Group)) +
  geom_point(size = 3) +
  geom_text(aes(label = Country.Name), vjust = -1, size = 3, show.legend = FALSE) +
  scale_color_manual(values = c("Countries with Highest Production" = "darkgreen", "Countries with Lowest Production" = "darkorange")) +
  labs(
    title = "CO"[2] ~ "Emissions over Countries with Lowest and Highest Livestock Production",
    x = "Total Livestock Production",
    y = "Total CO2 Emissions (Mt)",
    color = "Group"
  ) +
  theme(legend.position = "bottom")

```

Analyzing the Trend of CO<sub>2</sub> Emissions and Air Pollution over 30 Years
```{r}
emissions_df_pollution <- emissions_df

emissions_df_pollution$CO2_Emissions[is.na(emissions_df_pollution$CO2_Emissions)] <- 0
emissions_df_pollution$Air_Pollution[is.na(emissions_df_pollution$Air_Pollution)] <- 0

aggregated_data <- emissions_df_pollution |>
  group_by(Year) |>
  summarize(
    Total_CO2 = sum(CO2_Emissions),
    Total_Pollution = sum(Air_Pollution)
  )

aggregated_data <- aggregated_data |>
  mutate(
    Scaled_CO2 = (Total_CO2 - min(Total_CO2)) / (max(Total_CO2) - min(Total_CO2)),
    Scaled_Pollution = (Total_Pollution - min(Total_Pollution)) / (max(Total_Pollution) - min(Total_Pollution))
  )

ggplot(aggregated_data, aes(x = Year)) +
  geom_line(aes(y = Scaled_CO2, color = "CO2 Emissions (Mt)")) +
  geom_line(aes(y = Scaled_Pollution, color = "Air Pollution (µg/m³)")) +
  scale_color_manual(values = c("CO2 Emissions (Mt)" = "red", "Air Pollution (µg/m³)" = "darkblue")) +
  labs(
    title = "Worldwide CO"[2] ~ "Emissions and Air Pollution Trend (1990-2020)",
    x = "Year",
    y = "Scaled Values",
    color = "Legend"
  ) +
  theme(legend.position = "right")

```